### 2012-09-01 inst/examples/managed_warning.Rmd

``` {r }
pars <- c(r = .75, k = 10, a=1.7, H=1, Q = 3)
K <- 8 # approx
````

@@

``` {r miscpars}
delta <- 0.05
xT <- 0
OptTime <- 200
sigma_g <- .1
````

@@

``` {r  SDP_Mat}
SDP_Mat <- determine_SDP_matrix(f, pars, x_grid, h_grid, sigma_g)
````

@@

``` {r  simulate }
sims <- lapply(1:100, function(i){
  ForwardSimulate(f, pars, x_grid, h_grid, x0=K, opt$D, z_g, z_m, z_i)
})
````

## Summarize and plot the results

R makes it easy to work with this big replicate data set.  We make data tidy (melt), fast (data.tables), and nicely labeled.

``` {r  tidy}
dat <- melt(sims, id=names(sims[[1]]))  
dt <- data.table(dat)
setnames(dt, "L1", "reps") # names are nice
````

### Plots

``` {r  p0}
p0 <- ggplot(subset(dt,reps==1)) +
  geom_line(aes(time, fishstock)) +
  geom_abline(intercept=opt$S, slope = 0) +
  geom_line(aes(time, harvest), col="darkgreen") 
p0
````

This plot summarizes the stock dynamics by visualizing the replicates. Reed's S shown again, along with the dotted line showing the allee threshold, below which the stock will go to zero (unless rescued stochastically).

``` {r  p1}
p1 <- ggplot(dt) + geom_abline(intercept=opt$S, slope = 0) + 
  geom_abline(intercept=xT, slope = 0, lty=2) 
p1 <- p1 + geom_line(aes(time, fishstock, group = reps), alpha = 0.2)
p1
````

## Calculate warning signals on managed system

``` {r libs2}
library(earlywarning)
````

``` {r earlywarning}
acor_tau <- dt[, 
               warningtrend(data.frame(time=time, value=fishstock),
                            window_autocorr),
               by=reps]

var_tau <- dt[, 
              warningtrend(data.frame(time=time, value=fishstock),
                          window_var),
              by=reps]
````

``` {r modelbased}
m <- dt[, 
        stability_model(data.frame(time=time, value=fishstock),
                          "LSN")$pars["m"],
        by=reps]
````

``` {r summaryplot}
signals <- melt(data.frame(var=var_tau$V1, acor=acor_tau$V1, m=m$V1))
ggplot(signals) + geom_histogram(aes(value)) + facet_wrap(~variable, scales="free")
````

## Unmanaged warning signals
Calculate warning signals on the same set of shocks in unfished data

``` {r earlywarning2}
acor_tau <- dt[, 
               warningtrend(data.frame(time=time, value=unharvested),
                            window_autocorr),
               by=reps]

var_tau <- dt[, 
              warningtrend(data.frame(time=time, value=unharvested),
                          window_var),
              by=reps]
````

``` {r modelbased2}
m <- dt[, 
        stability_model(data.frame(time=time, value=unharvested),
                          "LSN")$pars["m"],
        by=reps]
````

``` {r summaryplot2}
signals <- melt(data.frame(var=var_tau$V1, acor=acor_tau$V1, m=m$V1))
ggplot(signals) + geom_histogram(aes(value)) + facet_wrap(~variable, scales="free")
````

>>>>

``` {r }
pars <- c(r = .75, k = 10, a=1, H=1, Q = 3)
K <- 8 # approx
````

@@

``` {r miscpars}
delta <- 0.005
xT <- 0
OptTime <- 5000
sigma_g <- .1
````

@@

``` {r  SDP_Mat}
pdfn <- function(P, s){
  dunif(P, 1 - s, 1 + s)
}
SDP_Mat <- determine_SDP_matrix(f, pars, x_grid, h_grid, sigma_g, pdfn)
````

@@

``` {r simulateme}
Dt <- 1

ForwardSimulate <- 
  function(f, pars, x_grid, h_grid, x0, D, z_g,
         z_m=function(x) 1, z_i = function(x) 1, 
         profit=NULL){

  OptTime <- dim(D)[2]    # Stopping time
  x_h <- numeric(OptTime) # population dynamics with harvest
  h <- numeric(OptTime) # optimal havest level
  x_h[1] <- x0  # initial values
  s <- x_h # also track escapement
  x <- x_h # What would happen with no havest
  p <- numeric(OptTime)

  ## Simulate through time ##
  for(t in 1:(OptTime-1)){

    # Move towards collapse
    pars[3] <- pars[3] + Dt/OptTime

    # Assess stock, with potential measurement error
    m_t <- x_h[t] * z_m()
    # Current state (is closest to which grid posititon) 
    St <- which.min(abs(x_grid - m_t)) 
    q_t <- h_grid[D[St, (t + 1) ]] 
    # Implement harvest/(effort) based on quota with noise 
    h[t] <- q_t * z_i()
    # Noise in growth 
    z <- z_g() 
    # population grows
    x_h[t+1] <- z * f(x_h[t], h[t], pars) # with havest
    s[t+1]   <- x_h[t] - q_t # anticipated escapement
    x[t+1]   <- z * f(x[t], 0, pars) # havest-free dynamics
    p[t] <- profit(x_h[t], h[t])
  }
  data.frame(time = 1:OptTime, fishstock = x_h, harvest = h,
             unharvested = x, escapement = s, profit = p) 
}

dat <- ForwardSimulate(f, pars, x_grid, h_grid, x0=K, opt$D, z_g, z_m, z_i)
x <- dat$fishstock
```

### Plot of timeseries

``` {r  p0}
plot(x, type='l')
````

Truncate the timeseries

``` {r truncate}
y <- x[x > 1.5]
````

``` {r  p1}
plot(y, type='l')
````

### Calculate warning signals on the truncated series.

``` {r libs2}
library(earlywarning)
dat <- data.frame(time=1:length(y), value=y)
````

``` {r earlywarning}
acor_tau <- warningtrend(dat, window_autocorr)
var_tau <- warningtrend(dat, window_var)

acor_tau
var_tau
````

## Model based statistics

Fit the models

``` {r modelbased}
A <- stability_model(dat, "OU")
B <- stability_model(dat, "LSN")
observed <- -2 * (logLik(A) - logLik(B))
m <- B$pars["m"]

observed
m
````

Compute summary stat versions

``` {r summarystat_roc} 
summarystat_roc <- function(A,B, summarystat_functions, reps=200){
  require(plyr)
  require(reshape2)
  Asim <- simulate(A, reps)
  Bsim <- simulate(B, reps)
  Asim <- melt(Asim, id = "time")
  Bsim <- melt(Bsim, id = "time")
  names(Asim)[2] <- "rep"
  names(Bsim)[2] <- "rep"
	dat <- lapply(summarystat_functions, function(f){
	  wsA <- ddply(Asim, "rep", warningtrend, f)
  	wsB <- ddply(Bsim, "rep", warningtrend, f)
	  data.frame(null = wsA$tau, test = wsB$tau)
	})
	tidy <- melt(dat)
}

dat <- summarystat_roc(A,B, list(var=window_var, acor=window_autocorr))
ggplot(dat) + geom_density(aes(value, fill=variable)) + facet_wrap(~L1)
```

Set up a parallel environment

``` {r parallel}
require(snowfall)
sfInit(par=T, cpu=12)
sfLibrary(earlywarning)
sfExportAll() 
```

Evaluate the ROC curve

``` {r roc}
reps <- sfLapply(1:500, function(i) compare(A, B))
lr <- lik_ratios(reps)
roc <- roc_data(lr)
```

Plot results.

``` {r plotroc}
require(ggplot2)
ggplot(lr) + geom_density(aes(value, fill = simulation), alpha = 0.6) + 
    geom_vline(aes(xintercept = observed))
ggplot(roc) + geom_line(aes(False.positives, True.positives))
```
